Uncertainty-aware graph-based multimodal remote sensing classification

### 2021-07-09 latex/CEUR/paper.tex
.pdf back within 6 pages again

The framework described in this section is the same as is described by the work of \cite{ZhaoXujiang2020} where they use the approach to perform semi-supervised classification of benchmark graph classification datasets including Cora and Pubmed. The method they present has been adapted here in order to perform the task of multimodal remote sensing classification.

@@

\subsection{Convergence assistance techniques}
In order to assist the convergence of subjective models, two additional assistance techniques have been used: teacher knowledge distillation and the use of a Dirichlet prior.
These have been shown to allow subjective models to provide better uncertainty estimates \cite{ZhaoXujiang2020}.
\subsubsection{Teacher knowledge distillation}
By training a non-subjective model in advance, its outputs, $\hat{p}_{ik}$, can be used in order to encourage the subjective model to converge to node Dirichlet distributions with $\mathbb{E}[p_{ik}]$ which are close to the teacher's deterministic estimates.
This is achieved using an additional term in the loss function,
\begin{equation}
    \mathcal{L}_{\text{T}}(\boldsymbol{\theta})=\sum_i\sum_k\left(\hat{p}_{ik}\log\frac{\hat{p}_{ik}}{\mathbb{E}[p_{ik}]}\right),
\end{equation}
which corresponds to the summation of KL-divergence terms between the teacher output probability and the expected value of the subjective model's Dirichlet distribution for each node, $\sum_i\infdiv{\hat{p}_{ik}}{\mathbb{E}[p_{ik}]}$.
Notice that this sum is computed over all nodes as opposed to just the nodes in $\mathbb{L}$.
Models trained using a teacher are denoted using the ``-T'' suffix e.g. a S-BGCN-T model would indicate that a pre-trained GCN was used as a teacher in order to assist the training convergence of a subjective graph convolutional model.
\subsubsection{Dirichlet prior}
A second convergence assistance technique which can be used involves the use of a Dirichlet prior, $\hat{\boldsymbol{\alpha}}$.
The exact method chosen to provide $\hat{\boldsymbol{\alpha}}$ will depend on the nature of the problem but we will assume nodes which are nearby in the graph are more likely to belong to the same output class than nodes which are far apart, a property known as homophily \cite{HuangQian2019}.
Using this assumption, we can use the computed distances on the graph in order to assign contributions of evidence from observed node labels to the other nodes in the graph using a function of our choosing.
If $d_{ij}$ denotes the shortest path distance between a given node, indexed by $i$ and an observed node, indexed by $j$, then the amount of evidence contributed to suggest that the $i^\text{th}$ node belongs to the $k^\text{th}$ class is given by
\begin{equation}
    h_{ik}(y_j,d_{ij})=
    \begin{cases}
    \frac{1}{\sqrt{2\pi\sigma^2}}\exp{\left(\frac{-d_{ij}^2}{2\sigma^2}\right)}, & \text{if } y_{jk}=1,\\
    0, & \text{otherwise},
    \end{cases}
\end{equation}
where $\sigma$ is a scale parameter which controls the order of distance magnitude over which evidence will propagate in the prior.
The total evidence to suggest the $i^\text{th}$ node belongs to the $k^\text{th}$ class, $e_{ik}$ can be found by summing these contributions over the nodes in the training set, such that the element in the prior is given by 
\begin{equation}
    \hat{\boldsymbol{\alpha}}_{ik}=1+e_{ik}=1+\sum_{j\in\mathbb{L}}h_{ik}(y_j, d_{ij}).
\end{equation}
The KL-divergence between the Dirichlet distribution of the prior and the model output is given by the term
\begin{equation}
    \mathcal{L}_{\text{K}}(\boldsymbol{\theta})=\sum_i\infdiv{\text{Dir}(\mathbf{p}_i\vert\boldsymbol{\alpha}_i)}{\text{Dir}(\hat{\mathbf{p}}_i\vert\hat{\boldsymbol{\alpha}}_i)},
\label{eq::teacher_loss}
\end{equation}
which can, in turn, be incorporated into the total loss function.
Models trained using a prior are denoted using the ``-K'' suffix.

@@

\begin{table}
\scriptsize
\renewcommand{\arraystretch}{1.5}
\caption{Loss function components and their weighting coefficients for different model types}
\label{loss_table}
\centering
\begin{tabular}{cc}
\hline
\bfseries Model name & \bfseries $\mathcal{L}_{\text{total}}(\boldsymbol{\theta})$\\
\hline
S-BGCN & $\mathcal{L}(\boldsymbol{\theta})$ \\
S-BGCN-T & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{T}}\mathcal{L}_{\text{T}}(\boldsymbol{\theta})$ \\
S-BGCN-K & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{K}}\mathcal{L}_{\text{K}}(\boldsymbol{\theta})$ \\
S-BGCN-T-K & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{T}}\mathcal{L}_{\text{T}}(\boldsymbol{\theta})+\lambda_{\text{K}}\mathcal{L}_{\text{K}}(\boldsymbol{\theta})$ \\
\hline
\end{tabular}
\end{table}

@@

\subsection{Network training and hyperparameters}
Models were implemented and trained using the TensorFlow library \cite{Abadi2015}.
In order to handle the imbalance of classes in the dataset, sample weighting was used.
Samples were given weights which were inversely proportional to the number of total samples of each class in the training set.
These weights such that the losses related to nodes belonging to under-represented classes were given an increased influence over parameter updates and vice versa.

@@

Each test was performed for different random dataset splits and model weight initializations in order to obtain mean and standard deviation measures of performance.

Benchmark models presented include the `vanilla' GCN (which can provide uncertainty via entropy) and the Drop-GCN which is a GCN where dropout inference has been applied in order to produce estimates of aleatoric and epistemic uncertainty in addition to entropy.

@@

Models of each type described (as well as benchmark models) were trained and their outputs analyzed in order to produce the AUROC and AUPR values seen in Table \ref{tab::misc_auroc}.

@@

The S-BGCN-T-K model trained achieves the highest ranked misclassification detection performance, with the uncertainty measured by entropy providing the best distinguishing metric, with AUROC and AUPR of $0.912$ and $0.672$ respectively, closely followed by the performance achieved by the measures of vacuity (AUROC and AUPR of $0.908$ and $0.659$ respectively) and aleatoric uncertainty (AUROC and AUPR of $0.905$ and $0.649$ respectively).

@@

The presented results have illustrated the importance of convergence assistance techniques as a means for improving the quality of uncertainty estimates, particularly through the use of a prior.
This can be seen by comparing the S-BGCN-T-K misclassification detection and OOD detection performance with equivalent models which do not use a prior, e.g. S-BGCN-T.
Of the uncertainty measures produced by the presented S-BGCN-T-K model, the measures of vacuity, aleatoric uncertainty or entropy all provide near-best performance for the task of identifying misclassifications.
In order to identify OOD nodes, the measure of vacuity stands out as the best-performing   uncertainty metric.

Future work should consider the generalisation potential of this method by considering its performance on other challenging remote sensing classification datasets.
There is also scope for research into how the choice of method for computing the $\hat{\boldsymbol{\alpha}}$ prior affects the quality of uncertainty estimates, either by varying the scale parameter, $\sigma$, or considering different prior computation methods entirely.
%% The acknowledgments section is defined using the "acknowledgments" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
% \begin{acknowledgments}
%   Thanks to ...
% \end{acknowledgments}

>>>>

The framework described in this section is the same as is described by the work of \cite{ZhaoXujiang2020} where they use the approach to perform semi-supervised classification of benchmark graph classification datasets including Cora and Pubmed.
The method they present has been adapted here to perform the task of multimodal remote sensing classification.

@@

\subsection{Convergence assistance techniques}
In order to assist the convergence of subjective models, two additional assistance techniques have been used: teacher knowledge distillation and the use of a Dirichlet prior.
These have been shown to allow subjective models to provide better uncertainty estimates \cite{ZhaoXujiang2020}.
\subsubsection{Teacher knowledge distillation}
By training a non-subjective model in advance, its outputs, $\hat{p}_{ik}$, can be used in order to encourage the subjective model to converge to node Dirichlet distributions with $\mathbb{E}[p_{ik}]$ which are close to the teacher's deterministic estimates.
This is achieved using an additional term in the loss function,
\begin{equation}
    \mathcal{L}_{\text{T}}(\boldsymbol{\theta})=\sum_i\sum_k\left(\hat{p}_{ik}\log\frac{\hat{p}_{ik}}{\mathbb{E}[p_{ik}]}\right),
\end{equation}
which corresponds to the summation of KL-divergence terms between the teacher output probability and the expected value of the subjective model's Dirichlet distribution for each node, $\sum_i\infdiv{\hat{p}_{ik}}{\mathbb{E}[p_{ik}]}$.
Notice that this sum is computed over all nodes as opposed to just the nodes in $\mathbb{L}$.
Models trained using a teacher are denoted using the ``-T'' suffix e.g. a S-BGCN-T model would indicate that a pre-trained GCN was used as a teacher in order to assist the training convergence of a subjective graph convolutional model.
\subsubsection{Dirichlet prior}
A second convergence assistance technique which can be used involves the use of a Dirichlet prior, $\hat{\boldsymbol{\alpha}}$.
The exact method chosen to provide $\hat{\boldsymbol{\alpha}}$ will depend on the nature of the problem but we will assume nodes which are nearby in the graph are more likely to belong to the same output class than nodes which are far apart, a property known as homophily \cite{HuangQian2019}.
Using this assumption, we can use the computed distances on the graph to assign contributions of evidence from observed node labels to the other nodes in the graph using a function of our choosing.
If $d_{ij}$ denotes the shortest path distance between a given node, indexed by $i$ and an observed node, indexed by $j$, then the amount of evidence contributed to suggest that the $i^\text{th}$ node belongs to the $k^\text{th}$ class is given by
\begin{equation}
    h_{ik}(y_j,d_{ij})=
    \begin{cases}
      \frac{\exp{\left(\frac{-d_{ij}^2}{2\sigma^2}\right)}}{\left(2\pi\sigma^2\right)^{1/2}}, & \text{if } y_{jk}=1,\\
    0, & \text{otherwise},
    \end{cases}
\end{equation}
where $\sigma$ is a scale parameter which controls the order of distance magnitude over which evidence will propagate in the prior.
The total evidence to suggest the $i^\text{th}$ node belongs to the $k^\text{th}$ class, $e_{ik}$ can be found by summing these contributions over the nodes in the training set, such that the element in the prior is given by 
\begin{equation}
    \hat{\boldsymbol{\alpha}}_{ik}=1+e_{ik}=1+\sum_{j\in\mathbb{L}}h_{ik}(y_j, d_{ij}).
\end{equation}
The KL-divergence between the Dirichlet distribution of the prior and the model output is given by the term
\begin{equation}
    \mathcal{L}_{\text{K}}(\boldsymbol{\theta})=\sum_i\infdiv{\text{Dir}(\mathbf{p}_i\vert\boldsymbol{\alpha}_i)}{\text{Dir}(\hat{\mathbf{p}}_i\vert\hat{\boldsymbol{\alpha}}_i)},
\label{eq::teacher_loss}
\end{equation}
which can, in turn, be incorporated into the total loss function.
Models trained using a prior are denoted using the ``-K'' suffix.

@@

\begin{table}[!t]
\scriptsize
\renewcommand{\arraystretch}{1.5}
\caption{Loss function components and their weighting coefficients for different model types}
\label{loss_table}
\centering
\begin{tabular}{cc}
\hline
\bfseries Model name & \bfseries $\mathcal{L}_{\text{total}}(\boldsymbol{\theta})$\\
\hline
S-BGCN & $\mathcal{L}(\boldsymbol{\theta})$ \\
S-BGCN-T & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{T}}\mathcal{L}_{\text{T}}(\boldsymbol{\theta})$ \\
S-BGCN-K & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{K}}\mathcal{L}_{\text{K}}(\boldsymbol{\theta})$ \\
S-BGCN-T-K & $\mathcal{L}(\boldsymbol{\theta})+\lambda_{\text{T}}\mathcal{L}_{\text{T}}(\boldsymbol{\theta})+\lambda_{\text{K}}\mathcal{L}_{\text{K}}(\boldsymbol{\theta})$ \\
\hline
\end{tabular}
\end{table}

@@

\subsection{Network training and hyperparameters}
Models were implemented and trained using the TensorFlow library \cite{Abadi2015}.
In order to handle the imbalance of classes in the dataset, sample weighting was used.
Samples were given weights which were inversely proportional to the number of total samples of each class in the training set.
This allows the losses related to nodes from under-represented classes to have an increased influence over parameter updates and vice versa.

@@

Each test was performed for different random dataset splits and model weight initializations to obtain mean and standard deviation measures of performance.

Benchmark models presented include the `vanilla' GCN (which can provide uncertainty via entropy) and the Drop-GCN which is a GCN where dropout inference has been applied to produce estimates of aleatoric and epistemic uncertainty in addition to entropy.

@@

Models of each type described (as well as benchmark models) were trained and their outputs analyzed to produce the AUROC and AUPR values seen in Table \ref{tab::misc_auroc}.

@@

The S-BGCN-T-K model achieves the best misclassification detection performance, with the uncertainty measured by entropy providing the best distinguishing metric, with AUROC and AUPR of $0.912$ and $0.672$ respectively, closely followed by the performance achieved by the measures of vacuity (AUROC and AUPR of $0.908$ and $0.659$ respectively) and aleatoric uncertainty (AUROC and AUPR of $0.905$ and $0.649$ respectively).

@@

The presented results illustrate the importance of convergence assistance techniques as a means for improving the quality of uncertainty estimates, particularly through the use of a prior.
This can be seen by comparing the S-BGCN-T-K misclassification detection and OOD detection performance with equivalent models which do not use a prior, e.g. S-BGCN-T.
Of the uncertainty measures produced by the S-BGCN-T-K model, the measures of vacuity, aleatoric uncertainty or entropy all provide near-best performance for the identifying misclassifications.
In order to identify OOD nodes, the measure of vacuity stands out as the best-performing uncertainty metric.

Future work should consider the generalisation potential of this method by assessing performance on other challenging remote sensing classification datasets.
There is also scope for research into how the choice of method for computing the $\hat{\boldsymbol{\alpha}}$ prior affects the quality of uncertainty estimates, either by varying the scale parameter, $\sigma$, or considering different prior computation methods entirely.
%% The acknowledgments section is defined using the "acknowledgments" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
% \begin{acknowledgments}
%   Thanks to ...
% \end{acknowledgments}
